{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-06-04 review_response.Rnw

I think that the caveats described above have been acknowledged by the authors, such as starting on line 527, but I would recommend framing the issues associated with the “end member” problem in a very straightforward way.

{\it We suggest that the model can work well in any tidal regime provided that the correlation between tidal changes and the solar cycle is low at the time scale of interest.  As we acknowledged, this condition was often not met in the case of Elkhorn Slough, leading to significant problems in applying the open water method with a single DO time series regardless of the method used.  Critically, it is possible to know by analyzing the correlation of tides and the solar cycle that our method would not be a good choice before applying it to Elkhorn Slough.  The 2014 paper by Nidzieko et al. illustrated the complex and interesting dynamics there and showed that they are better explored with a more detailed data set than our method is intended to require.  It is useful to point out that simply ignoring advection is also not a good option in estuaries like Elkhorn Slough, unless perhaps you are only interested in the means at the annual or longer time scale.

Our method works by statistically separating DO variability associated with tide height from DO variability associated with the solar cycle.  We think that it is unlikely that we would have obtained the results that we did by generating offsetting errors.

Although we mentioned these points previously (e.g., lines 431, 521) we added additional explanation based on the reviewer’s comments. Specifically, we added text to the comments and recommendations section (starting on line 527) to emphasize how and why the model can effectively characterize and remove the tidal signal, emphasizing that the model can work well in a variety of tidal regimes.

@@

{\it We recognize that tide height is not the same measure as advection, and specifically, that advective currents can vary in both magnitude and phase relative to tide height.  Equation 13 describes our method for computing a simulated component of the DO time series associated with advection.  This equation describes the DO changes, not the magnitude of the tidal current as the reviewer’s question suggests.  We decided that equation 13 actually provided excessive detail which confused more than it informed, and therefore we eliminated the equation from the text.  The correct interpretation is that the DOadv ``end members'' occur at either high or low tide when the water mass and associated DO gradient is maximally displaced either ``up-estuary'' or ``down-estuary.'' Our approach assumes that this occurs around high tide and low tide, respectively, such that DOadv is linearly related to H (i.e., our eq. 1). Because of this, our model is most applicable in the case of a standing wave, which is reasonable for the types of estuarine sites where this approach is most applicable.

Fortunately, our weighted regression method only requires a variable that is statistically correlated with DO changes associated with displacement of the water mass by advection.  We need not predict either the actual current speeds or the length of the tidal excursion, which we describe in eq. 10 and 11 only to conceptualize the processes that we model empirically.  We added additional text to make these points more clear, beginning on line 123. We also revised lines 204-213 in the methods to clarify differences between how we generated the simulated data and how what we expected in the actual time series.

Line 121: `Tidal height is used as a proxy for advection because this measurement is widely available, easy to measure, and meets the model requirements, which is a variable that is correlated locally in time with DO variations attributable to advection.  Importantly, tide height need not predict the actual magnitude of advection.  This is important because, across estuarine locations, the relationship between tide height and tidal currents can be variable in both magnitude and phase, depending on local characteristics.  The weighted regression model is well-suited for the situation in estuaries because it empirically fits the local relationship between tide height and DO, such that no a priori knowledge of the relationship between tide height and either the magnitude of advection or the horizontal gradient in DO is needed.'

Line 204: `A tidal series was simulated based on the principal lunar semidiurnal (M2) tide with a period of 12.42 hours (Foreman and Henry 1989).  The amplitude was set to 1 meter and centered  at 4 meters.  The tidal time series simulated DO changes with advection, $DO_{adv}$ (eq. (7) and Fig. 1). Conceptually, this vector represents the rate of change in DO as a function of horizontal water movement from tidal advection such that:
\begin{equation} \tag{10}
\frac{\delta DO_{adv}}{\partial t} = \frac{\partial DO}{\partial x} \cdot \frac{\partial x}{\partial t}
\end{equation}
where the change in DO from advection over time $t$ is the product of the change in DO over the change in tidal excursion $x$ and the change in tidal excursion over time.  The simulated tidal time series was used to create $DO_{adv}$ as a simple sine wave as in eq. (8).  This time series was centered at zero to simulate increasing or decreasing DO with tidal height change.  The final time series for observed DO was the sum of biological DO and advection DO (eq. (4) and Fig. 1). For time series from the NERRS monitoring sites, changes in DO associated with advection are modeled as linear function of tidal height as indicated by the parameter $\beta_{2}$ in eq.(1).  We can express the tidal component of DO variations alone as
\begin{equation} \tag{11}
DO_{adv} = \beta_{2}H
\end{equation}'
}

### 2015-06-04 assess.Rnw
proofed word version and review to make sure I included everything, sent to Jim

Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.  The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).  Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.  Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
\begin{equation} \label{do_unc_n}
DO_{unc, n} = \varepsilon_{obs, n} + \sum_{t = 1}^{n} \varepsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\varepsilon_{obs}$, $\varepsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.  The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).

@@

The observed and filtered daily estimates were averaged by month for each site to evaluate the effect of filtering on mean production and respiration at the monthly time scale (\cref{tab:case_res,fig:metab_sum}).  Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.  Results for Sapelo Island suggested that time averaged estimates, either as monthly or annual means, were comparatively unchanged by filtering, although there were some more subtle differences seasonally.  Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering. It is likely that both the filtered and unfiltered estimates are erroneous; the result of collinearity between the tidal and solar cycle.  Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.  Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales (\cref{tab:case_res}).

>>>>

Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.  The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).  Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to an ecosystem, it is sufficient to address the analytical objectives of the simulations.  Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
\begin{equation} \label{do_unc_n}
DO_{unc, n} = \varepsilon_{obs, n} + \sum_{t = 1}^{n} \varepsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\varepsilon_{obs}$, $\varepsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.  The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).

@@

The observed and filtered daily estimates were averaged by month for each site to evaluate the effect of filtering on mean production and respiration at the monthly time scale (\cref{tab:case_res,fig:metab_sum}).  Filtered production and respiration estimates for Padilla Bay and Rookery Bay exhibited monthly variation that was more characteristic of expected trends during warmer months than in the unfiltered estimates.  Results for Sapelo Island suggested that time averaged estimates, either as monthly or annual means, were comparatively unchanged by filtering, although there were some more subtle differences seasonally.  Results for Elkhorn Slough varied significantly such that summer production and respiration were significantly reduced by filtering. It is likely that both the filtered and unfiltered estimates are erroneous; the result of collinearity between the tidal and solar cycle.  Overall, these trends emphasize the importance of considering different averaging periods for interpreting metabolism estimates.  Each case study showed differences in observed and filtered values for monthly means, suggesting tidal variation may influence metabolism estimates at relatively long time scales.
